### 2019-02-20 2-Chapters/3-Chapter/chapter3.tex
Added a lot of content for Chapter 3 on SMPyBandits.

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/3-Chapter/Images/}}

@@

@@

If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and ensure XXX.

\paragraph{Simulation loop for multi-player bandit}
%
We refer to Chapter~\ref{chapter:5} which introduces and details various multi-player bandit models.
Any simulation library targeting multi-player bandit problems must implement at least another component:
a collision model, and a simulation loop that essentially looks like this.
\begin{itemize}
    \item First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask each player $\cA^{(j)}$ his chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment. The
    sample (random) feedback $r_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $Y^{(j)}(t)$ from the random feedback and the collision model. Finally feeds the observation $(A^{(j)}(t),Y^{(j)}(t))$ to the $M$ players,
    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
\end{itemize}

\paragraph{Reward distributions}
%
This library tackles one dimensional distributions,
and supports [`Bernoulli`](https://smpybandits.github.io/docs/Arms.Bernoulli.html), [`binomial`](https://smpybandits.github.io/docs/Arms.Binomial.html), [`Poisson`](https://smpybandits.github.io/docs/Arms.Poisson.html), and a generic [`discrete`](https://smpybandits.github.io/docs/Arms.DiscreteArm.html) distributions,
as well as [`exponential`](https://smpybandits.github.io/docs/Arms.Exponential.html), [`gamma`](https://smpybandits.github.io/docs/Arms.Gamma.html), [`Gaussian`](https://smpybandits.github.io/docs/Arms.Gaussian.html) and [`uniform`](https://smpybandits.github.io/docs/Arms.Uniform.html) continuous distributions,
which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).

SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
containing [over 65](https://SMPyBandits.GitHub.io/docs/Policies.html) algorithms.
It uses a well-designed hierarchical structure and [class inheritance scheme](https://SMPyBandits.GitHub.io/uml_diagrams/README.html) to minimize redundancy in the codebase.
For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the [`IndexPolicy`](https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html) class.
An index-based algorithm computes an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$ and simply play $A(t) = \arg\max_k I_k(t)$.

We use a well-designed hierarchical structure and [class inheritance scheme](https://smpybandits.github.io/uml_diagrams/README.html) to minimize redundancy in the codebase, and for instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented), by inheriting from the [`IndexPolicy`](https://smpybandits.github.io/docs/Policies.IndexPolicy.html) class:

@@

### Multi-Players MAB

For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the *same* $K$ arms. Whenever two or more players select the same arm at the same time, they all suffer from a collision.
Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (*i.e.*, arms with highest means) without collisions.
*SMPyBandits* implements [all collision models](https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html) found in the literature, as well as all the algorithms from the last 10 years (including [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html), [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html), and our state-of-the-art algorithms [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) from @BessonALT2018).
For comparison, realistic or full-knowledge centralized algorithms are also implemented.

---

@@

A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: [`SMPyBandits.GitHub.io`](https://SMPyBandits.GitHub.io/).

@@

We show how to install SMPyBandits, and an example of how to run a simple experiment.
This bash snippet\footnote{~See [`SMPyBandits.GitHub.io/How_to_run_the_code.html`](https://SMPyBandits.GitHub.io/How_to_run_the_code.html) for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see [`pypi.org/project/SMPyBandits`](https://pypi.org/project/SMPyBandits/). You can install it directly with `sudo pip install SMPyBandits`, or from a `virtualenv` \cite{virtualenv}},
and install the requirements for Python 3 (once):

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
# 1. get the code in the folder you want
$ git clone https://GitHub.com/SMPyBandits/SMPyBandits.git
$ cd SMPyBandits.git
# 2. install the requirements
$ pip install -r requirements.txt
        }
        \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
        % \captionof{listing}{Small snippet of Bash to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
        \label{lst:3:howToInstallLibrary}
    \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
# 3. run a single-player simulation
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
    MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] python3 main.py configuration.py
        }
        \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
        % \captionof{listing}{Small snippet of Bash to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
        \label{lst:3:howToRunBasicLibrary}
    \end{listing}
\end{small}

### Example of a simulation and illustration

A small script [`configuration.py`](https://SMPyBandits.GitHub.io/docs/configuration.html) is used to import the [arm classes](https://SMPyBandits.GitHub.io/docs/Arms.html), the [policy classes](https://SMPyBandits.GitHub.io/docs/Policies.html) and define the problems and the experiments.
Choosing the algorithms is easy by customizing the `configuration["policies"]` list in the `configuration.py` file.
For instance, one can compare the standard anytime [`klUCB`](https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html) algorithm against the non-anytime variant [`klUCBPlusPlus`](https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html) algorithm, and also [`UCB`](https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html) (with $\alpha=1$) and [`Thompson`](https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html) (with [Beta posterior](https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html)).

```python
configuration["policies"] = [
  { "archtype": klUCB, "params": { "klucb": klucbBern } },
  { "archtype": klUCBPlusPlus, "params": { "horizon": HORIZON, "klucb": klucbBern } },
  { "archtype": UCBalpha, "params": { "alpha": 1 } },
  { "archtype": Thompson, "params": { "posterior": Beta } }
]
```

Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in [a HDF5 file](http://docs.h5py.org/en/stable/high/file.html)\footnote{~For example, this simulation produces this HDF5 file, [GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5](https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5).}.
Figure \ref{fig:plot1} below shows the average regret for these $4$ algorithms.
The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.

@@

@@

SMPyBandits was used for the following research articles since $2017$:

- For \cite{Besson2018ALT}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/MultiPlayers.html`](https://SMPyBandits.GitHub.io/MultiPlayers.html) on the documentation.}. We designed the two [`RandTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html) and [`MCTopM`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html) algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (i.e., [`rhoRand`](https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html), [`MEGA`](https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html) and [`MusicalChair`](https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html)).

- In \cite{Besson2018WCNC}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page [`SMPyBandits.GitHub.io/Aggregation.html`](https://SMPyBandits.GitHub.io/Aggregation.html) on the documentation.}. We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called [`Aggregator`](https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html). Aggregating experts is a well-studied idea in sequential learning and in machine learning in general. We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts. This idea and algorithm can have interesting impact for Opportunistic Spectrum Access applications \cite{Jouini09} that use multi-armed bandits algorithms for sequential learning and network efficiency optimization.

- In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page [`SMPyBandits.GitHub.io/DoublingTrick.html`](https://SMPyBandits.GitHub.io/DoublingTrick.html) on the documentation.}. In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments. A well-known trick for transforming any non-anytime algorithm to an anytime variant is the "Doubling Trick": start with an horizon $T_0\in\mathbb{N}^$, and when $t > T_i$, use $T_{i+1} = 2 T_i$. We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results. A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$). And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting. It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).

- In \cite{Besson2019GLRT}, we used SMPyBandits for piecewise stationary MAB models (single-player). We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page [`SMPyBandits.GitHub.io/NonStationary.html`](https://SMPyBandits.GitHub.io/NonStationary.html) on the documentation.}. XXX

@@

Below is presented in Listing~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
Only the mean reward is printed at the end of the simulation.
%
With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^ = \mu_2$.
\begin{verbatim}
The UCB algorithm obtains here a mean reward = 0.886
\end{verbatim}

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Example of use of SMPyBandits \label{lst:3:exampleOfUse}.}
    % \label{lst:3:exampleOfUse}
% \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Small example of \texttt{main.py} file\label{lst:3:main}.}
    % \label{lst:3:main}
% \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \minted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}{
$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
        }
        \caption{Small snippet of Bash to run an experiment}
        % \captionof{listing}{Small snippet of Bash to run an experiment \label{lst:3:howToRunExperiment2}.}
        \label{lst:3:howToRunExperiment2}
    \end{listing}
\end{small}

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
    % \caption{Small example of \texttt{configuration.py} file}
    \captionof{listing}{Small example of \texttt{configuration.py} file\label{lst:3:configuration}.}
    % \label{lst:3:configuration}
% \end{listing}
\end{small}

@@

Quickly explain the advantages and drawbacks of using Joblib and joblib.Parallel to run simulations on multiple cores of the same computer

>>>>

This Chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

\paragraph{Publications}
%
This Chapter is based on the following publication: \cite{SMPyBanditsJMLR}.
The code code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is publicly released under the MIT open-source License.

@@

% FIXME finish this paragraph
If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(r_k(t))_{k\in[K], t\in[T]}$ before-hand, and ensure XXX.

\paragraph{Simulation loop for multi-player bandit}

For Cognitive Radio and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
Whenever two or more players select the same arm at the same time, they all suffer from a collision.
%
Different collision models has been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
We refer to Chapter~\ref{chapter:5} which introduces and details various multi-player bandit models.

SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms from the last 10 years (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}).
%
It includes
\texttt{rhoRand} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}}) from \cite{Anandkumar11},
\texttt{MEGA} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}}) \cite{Avner15},
\texttt{MusicalChair} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}}) from \cite{Rosenski16},
and our state-of-the-art algorithms
\texttt{RandTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}}) and
\texttt{MCTopM} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}}) from \cite{Besson2018ALT}.
For comparison, realistic or full-knowledge centralized algorithms are also implemented.

Any simulation library targeting multi-player bandit problems must implement at least another component:
a collision model, and a simulation loop that essentially looks like this.
\begin{itemize}
    \item First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    \item Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask each player $\cA^{(j)}$ his chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment. The
    sample (random) feedback $r_k(t)$ (\iid) from distributions $\nu_{k}$, and compute the rewards $Y^{(j)}(t)$ from the random feedback and the collision model. Finally feeds the observation $(A^{(j)}(t),Y^{(j)}(t))$ to the $M$ players,
    \item At the end, compute the accumulated reward, the regret, plot visualizations etc.
\end{itemize}

\subsection{Reward distributions}
%
This library tackles one dimensional distributions,
and supports \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Arms.Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Arms.Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Arms.Poisson}}), and a generic \emph{discrete} (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{Arms.DiscreteArm}}) distributions,
as well as \emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Arms.Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Arms.Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Arms.Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Arms.Uniform}}) continuous distributions,
which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$).

The default is to use the same distribution for the $K$ arms, but it also possible to mix them.

% FIXME
\TODOL{point to the notebook!}
% https://perso.crans.org/besson/phd/notebooks/Easily_creating_MAB_problems.html

\subsection{MAB algorithms}

SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
containing over 65 (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}) algorithms.
It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams showed on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
For instance, most existing algorithms are index-based, and new ones can be written easily by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
An index-based algorithm computes an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$ and simply play $A(t) = \arg\max_k I_k(t)$.
For instance the code specific to the UCB algorithm \cite{LaiRobbins85,Auer02} is as short as this (and fully documented):

@@

@@

A complete Sphinx documentation, for each algorithm and all parts of the codebase, even including the constants in the different configuration files, is available here: \texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.

@@

We show how to install SMPyBandits, and an example of how to run a simple experiment.
This bash snippet\footnote{~See this page \texttt{\href{https://SMPyBandits.GitHub.io/How_to_run_the_code.html}{SMPyBandits.GitHub.io/How\_to\_run\_the\_code.html}} for more details.} shows how to clone the code\footnote{~SMPyBandits is also available on Pypi, see \texttt{\href{https://pypi.org/project/SMPyBandits/}{pypi.org/project/SMPyBandits}}. You can install it directly with \texttt{sudo pip install SMPyBandits}, or from a \texttt{virtualenv} \cite{virtualenv}.},
and install the requirements for Python 3 (once):

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 1. get the code in the folder you want
$ git clone https://GitHub.com/SMPyBandits/SMPyBandits.git
$ cd SMPyBandits.git
# 2. install the requirements
$ pip install -r requirements.txt
        \end{minted}
        \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
        % \captionof{listing}{Small snippet of Bash to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
        \label{lst:3:howToInstallLibrary}
    \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 3. run a single-player simulation
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] \
python3 main.py configuration.py
    \end{minted}
    \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
    % \captionof{listing}{Small snippet of Bash to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
    \label{lst:3:howToRunBasicLibrary}
\end{listing}

\subsection{Example of a simulation and illustration}

A small script \texttt{configuration.py} (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}}) is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{SMPyBandits.GitHub.io/docs/Arms.html}}), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{SMPyBandits.GitHub.io/docs/Policies.html}}) and define the problems and the experiments.
Choosing the algorithms is easy by customizing the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{Policies.klUCB}}) against the non-anytime variant $\klUCB^{++}$ algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{Policies.klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{Policies.UCB}}) and Thompson sampling (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Policies.Thompson}}) with a Beta posterior (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Policies.Posterior.Beta}})).

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}
configuration["policies"] = [
  {"archtype": klUCB, "params": {"klucb": klucbBern}},
  {"archtype": klUCBPlusPlus, "params": {"horizon": HORIZON, "klucb": klucbBern}},
  {"archtype": UCBalpha, "params": {"alpha": 1}},
  {"archtype": Thompson, "params": {"posterior": Beta}}
]
    \end{minted}
    \caption{Small snippet of Python code to configure the list of algorithms tested on a problem.}
    % \captionof{listing}{Small snippet of Python code to configure the list of algorithms tested on a problem. \label{lst:3:howToConfigureAlgorithms}.}
    \label{lst:3:howToConfigureAlgorithms}
\end{listing}

Running the simulation as shown above will save figures in a sub-folder, as well as save data (pulls, rewards and regret) in a HDF5 file (\cite{h5py}, \texttt{\href{http://docs.h5py.org/en/stable/high/file.html}{docs.h5py.org/en/stable/high/file.html}})\footnote{~For example, this simulation produces this HDF5 file \texttt{\href{https://github.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}{GitHub.com/SMPyBandits/SMPyBandits/blob/master/plots/paper/example.hdf5}}.}.
Figure~\ref{fig:3:firstPlot} below shows the average regret for these $4$ algorithms.
The regret is the difference between the cumulated rewards of the best fixed-armed strategy (which is the oracle strategy for stationary bandits), and the cumulated rewards of the considered algorithms.
The Figure~\ref{fig:3:firstPlot_hist} shows the histogram of regret for the same experiment.

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.95\linewidth]{3.pdf}
	\caption[Example of a single-player simulation showing the average regret and histogram of regrets of $4$ algorithms]{
		Example of a single-player simulation showing the average regret and histogram of regrets of four algorithms. They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proved to match the lower-bound up-to a constant), and each but UCB is known to be optimal (\ie, with the constant matching the lower-bound). For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).
	}
	\label{fig:3:firstPlot}
\end{figure}

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.95\linewidth]{3_hist.pdf}
	\caption{Histogram of regret for the same experiment.}
	\label{fig:3:firstPlot_hist}
\end{figure}

@@

All the quoted libraries are free and open-source and can be installed in one command using the \texttt{pip} (\texttt{\href{https://pip.pypa.io/}{pip.PyPa.io}}) or \texttt{conda} (\texttt{\href{http://conda.anaconda.org/}{conda.anaconda.org}}) package manager.

@@

SMPyBandits was used for the following research articles since $2017$.

In \cite{Besson2018WCNC} (see Chapter~\ref{chapter:2}), we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation.}. We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}. Aggregating experts is a well-studied idea in sequential learning and in machine learning in general. We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts. This idea and algorithm can have interesting impact for Opportunistic Spectrum Access applications \cite{Jouini09} that use multi-armed bandits algorithms for sequential learning and network efficiency optimization.

- For \cite{Besson2018ALT} (see Chapter~\ref{chapter:5}), we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}. We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).

In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}. In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments. A well-known trick for transforming any non-anytime algorithm to an anytime variant is the "Doubling Trick": start with an horizon $T_0\in\mathbb{N}^$, and when $t > T_i$, use $T_{i+1} = 2 T_i$. We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results. A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$). And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting. It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).

In \cite{Besson2019GLRT} (see Chapter~\ref{chapter:6}), we used SMPyBandits for piece-wise stationary MAB models (single-player). We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}. XXX

\TODOL{Write a better explanation of this last paper.}

@@

Below is presented in Listing~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
Only the mean reward is printed at the end of the simulation.
%
With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
\begin{verbatim}
The UCB algorithm obtains here a mean reward = 0.886
\end{verbatim}

% https://tex.stackexchange.com/a/12430/
% \begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
    % \label{lst:3:exampleOfUse}
% \end{listing}
% \end{small}

@@

% https://tex.stackexchange.com/a/12430/
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Small example of \texttt{main.py} file\label{lst:3:main}.}
    % \label{lst:3:main}
% \end{listing}

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
    \end{minted}
    \caption{Small snippet of Bash to run an experiment}
    % \captionof{listing}{Small snippet of Bash to run an experiment \label{lst:3:howToRunExperiment2}.}
    \label{lst:3:howToRunExperiment2}
\end{listing}

% https://tex.stackexchange.com/a/12430/
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
    % \caption{Small example of \texttt{configuration.py} file}
    \captionof{listing}{Small example of \texttt{configuration.py} file\label{lst:3:configuration}.}
    % \label{lst:3:configuration}
% \end{listing}

@@

\TODOL{Write it!}
